From physical principles to ???

### 2016-02-07 2015CompleteDerivation/completeDerivation.tex
Adding physical distinguishability and topology part

\section{Determinism and reversibility}

@@

\begin{defn}\label{statedef}
The state space $\mathcal{S}$ of a physical system is a set.
\end{defn}

>>>>

\section{On studying a physical system}

@@

Before we are able to define how states should be captured mathematically, we need to better specify what we mean by physically distinguishable. Therefore we need to characterise what we mean by measurement, particularly because it has become such an overloaded term.

Consider the motion of a cannonball under inertial and gravitational forces. Light will be scattering off of it; as it lands the ground will be deformed and its temperature will rise. Those external physical process, which happen no matter what we do, can be used to distinguish the motion of the cannonball as their outcome are correlated. Therefore we can learn the position by looking at the reflected light, and learn the final kinetic momentum by looking at the deformation of the ground. In principle, any external process that has a correlation with the states under study can be used to perform a measurement, and any measurement is based on such a process. That is: for us a measurement is simply a physical process that we can use to distinguish states. There is no particular role played by the observer, except maybe create the device and forcing the system under study to interact with it.

This external process may be quite complicated: when a charged particle enters a calorimeter, a shower of particles is produced, photons are captured and are directed to photomultipliers, a current is read out, the current is then digitized, and so on. Some processes interfere with the system, they affect its dynamics, and other are destructive, the system no longer can be described by the original set of states. A bubble chamber is an example of the first (e.g. the magnetic field curves the motion of a charged particle); burning a substance to determine its caloric content is an example of the latter. In these cases, special care needs to make sure to link the outcomes with the original state, unaffected state. But in the end, the conceptual description remains the same.

We have a physical process that provides an outcome. This process has a set of potential outcomes. To each potential outcome we can associate a set of states. For example, if the electron follows this path, then its state is among the ones that has spin $1/2$. Since all measurements are finite precision, each outcome can be labelled by a finite set of digits, so the set of outcomes is countable. Note that different outcomes for the same process can overlap. For example, $4.12 \pm 0.05$ cm and $4.13 \pm 0.05$ cm are both legitimate possible outputs of the same measurement device. But since all states must be distinguishable, given two arbitrary states there must be a process precise enough to tell them apart. That is: the potential outcomes associated with the two states do not overlap. For example, $4.12 \pm 0.0005$ cm and $4.13 \pm 0.0005$ cm do not overlap anymore.

We can also combine two outcome of two different processes. For example, if one tells us spin is $+1/2$ and another that position is $4.13 \pm 0.05$ cm, then can say the state has both. That is, the state is inside both of the sets of states associated with the two outcomes: the intersection of two sets associated with an outcome is also a set associated with an outcome. We can make similar arguments about the union as we can always reduce the precision by combining different outcome of the same process.

This model maps very naturally to a topological space. The states are the elements of a set and the physical outcome provides a topology on that set. The fact that two elements of the set can be distinguished requires the space to be Hausdorff. The fact that the potential outcome of a process are countable requires the space to be second countable.

\begin{defn}\label{statedef}
The state space $\mathcal{S}$ of a physical system is a Hausdorff and second countable topological space.
\end{defn}
